---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# countmissingtool

<!-- badges: start -->

<!-- badges: end -->

The goal of `countmissing-tool` is to provide a simple and effective way to count missing values in all columns of a data frame, grouped by a specified variable. The `count_all_missing_by_group()` function returns a summary table with one row per group level, helping users assess missing data patterns across grouped subsets.

## Installation Instructions

You can install the development version of `countmissingtool` from [GitHub](https://github.com/) as follows:

1. If you haven't already, install `devtools` package: 

`
install.packages("devtools")
`

2. Install `countmissingtool` from Github in the R console:

`
devtools::install_github("stat545ubc-2024/countmissingtool", ref = "1.0.0")
`


## Example Usage

### Example 1: Count Missing Values by Group with different datasets
This example demonstrates how to count missing values in the R built-in datasets  `mtcars` and `airquality`, grouped by the `am` (transmission type) and `Month` (month of the year) columns, respectively.

``` {r}
library(countmissingtool) # To load the package

# Count missing values by the `am` column in the `mtcars` dataset
count_all_missing_by_group(mtcars, am)
```

``` {r}
# Count missing values by the `Month` column in `airquality`
count_all_missing_by_group(airquality, Month)
```

In both cases, the function will return a summary table as a tibble (dataframe) with one row for each group (transmission type or month), showing the count of missing values for each column in the dataset. This helps you quickly assess missing data patterns across different groups.

### Example 2: Handle Special Grouping Options
If you'd like to keep the grouping structure in the output, you can optionally specify the `.groups` argument. For instance, setting `.groups = "keep"` retains the grouping structure after summarizing the missing values.

``` {r}
# Count missing values by `Month` and keep grouping structure
count_all_missing_by_group(airquality, Month, .groups = "keep")
```

This ensures that the result is still grouped by `Month`, which might be useful if you plan to perform additional operations on the summarized data. 

There are also other possible actions available for handling grouping detailed in the function documentation, including: `"drop_last"`, "`drop"`, and `"rowwise"`. See the [countmissingtool.R file](https://github.com/stat545ubc-2024/assignment-b2-andreawongkoo/tree/b5d61e016d58da9863756bf89600c5edad3cd071/R) for more details.

