---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# countmissing

<!-- badges: start --> <!-- badges: end -->
The goal of `countmissing` is to provide a simple and effective way to count missing values in all columns of a data frame, grouped by a specified variable. The `count_all_missing_by_group()` function returns a summary table with one row per group level, helping users assess missing data patterns across grouped subsets.

## Installation
Follow these steps to install the development version of `countmissing` from GitHub:


### 1. Install devtools package if you haven't already

`if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}`

### 2. Install countmissing from GitHub
`devtools::install_github("andreawongkoo/countmissing", ref = "0.1.0")`

## Example

**Example 1:** This is a basic example that demonstrates how to use `count_all_missing_by_group()` to count missing values by group in the built-in `mtcars` R dataset:

``` {r}
library(countmissing)

# Example : Count missing values by the `am` column in mtcars
count_all_missing_by_group(mtcars, am)

```

The `count_all_missing_by_group()` function makes it easy to summarize missing values across grouped data. 

**Example 2:** For instance, you can use it to analyze missing data within specific categories, such as different months in the in-built `airquality` R dataset:

``` {r}

# Example: Count missing values by the `Month` column in airquality
count_all_missing_by_group(airquality, Month)
```
