---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# countmissing

<!-- badges: start -->

<!-- badges: end -->

The goal of `countmissing` is to provide a simple and effective way to count missing values in all columns of a data frame, grouped by a specified variable. The `count_all_missing_by_group()` function returns a summary table with one row per group level, helping users assess missing data patterns across grouped subsets.

## Installation

Follow these steps to install the development version of `countmissing` from GitHub:

### 1. Install devtools package if you haven't already

`if (!requireNamespace("devtools", quietly = TRUE)) {   install.packages("devtools") }`

### 2. Install countmissing from GitHub

`devtools::install_github("andreawongkoo/countmissing", ref = "0.1.0")`

## Example Usage

### Example 1: Count Missing Values in mtcars by Transmission Type
This example demonstrates how to count the missing values for each column of the R built-in `mtcars` dataset, grouped by the `am` column, which represents the type of transmission (automatic vs. manual).

``` {r}
library(countmissing)

# Count missing values by the `am` column in the `mtcars` dataset
count_all_missing_by_group(mtcars, am)
```

This will return a summary showing the count of missing values in each column for each transmission type (`am`), helping you quickly assess data quality across different groups.

### Example 2: Analyze Missing Data by Month in the airquality Dataset
You can also use the function to check missing data patterns in different subsets of a dataset. Here, we analyze the R built-in `airquality` dataset, grouping by the `Month` column.

``` {r}
# Count missing values by the `Month` column in `airquality`
count_all_missing_by_group(airquality, Month)
```

In this case, the function will summarize the missing values for each month, allowing you to understand potential patterns of missingness over time.

### Example 3: Handle Special Grouping Options
If you'd like to keep the grouping structure in the output, you can specify the `.groups` argument. For instance, setting .groups = "keep" retains the grouping structure after summarizing the missing values.

``` {r}
# Count missing values by `Month` and keep grouping structure
count_all_missing_by_group(airquality, Month, .groups = "keep")
```

This ensures that the result is still grouped by Month, which might be useful if you plan to perform additional operations on the summarized data.

